---
title: "R Refresher Activity Answers"
output: html_notebook
---

Load the mtcars dataset

```{r}
data(mtcars)
head(mtcars)
```

Problem 1: Descriptive Statistics

```{r}
# Summary statistics for mpg
mean_mpg <- mean(mtcars$mpg)
median_mpg <- median(mtcars$mpg)
range_mpg <- range(mtcars$mpg)
sd_mpg <- sd(mtcars$mpg)

# Print the results
cat("Mean of mpg:", mean_mpg, "\n")
cat("Median of mpg:", median_mpg, "\n")
cat("Range of mpg:", range_mpg, "\n")
cat("Standard Deviation of mpg:", sd_mpg, "\n")
```


Problem 2: Data Visualization

```{r}
# Histogram of mpg
hist(mtcars$mpg, main="Histogram of MPG", xlab="Miles per Gallon", col="lightblue", border="black")

# Scatter plot of mpg vs hp
plot(mtcars$hp, mtcars$mpg, main="MPG vs Horsepower", xlab="Horsepower", ylab="Miles per Gallon", pch=19, col="blue")
abline(lm(mtcars$mpg ~ mtcars$hp), col="red")
```

Problem 3: Correlation

```{r}
# Correlation between mpg and hp
cor_mpg_hp <- cor(mtcars$mpg, mtcars$hp)
cat("Correlation between mpg and hp:", cor_mpg_hp, "\n")
```

Problem 4: Linear Regression

```{r}
# Fit a linear model: mpg ~ hp
model <- lm(mpg ~ hp, data=mtcars)
summary(model)

# Plot the regression line
plot(mtcars$hp, mtcars$mpg, main="MPG vs Horsepower with Regression Line", xlab="Horsepower", ylab="Miles per Gallon", pch=19, col="blue")
abline(model, col="red")
```

Problem 5: Grouping and Aggregation

```{r}
# Group by number of cylinders and calculate average mpg
agg_mpg <- aggregate(mpg ~ cyl, data=mtcars, FUN=mean)
cat("Average MPG by Cylinder Count:\n")
print(agg_mpg)
```

Problem 6: Outliers Detection

```{r}
# Boxplots to identify outliers for multiple variables
par(mfrow=c(2, 2))
boxplot(mtcars$mpg, main="Boxplot for MPG", ylab="Miles per Gallon", col="lightblue")
boxplot(mtcars$hp, main="Boxplot for Horsepower", ylab="Horsepower", col="lightgreen")
boxplot(mtcars$wt, main="Boxplot for Weight", ylab="Weight (1000 lbs)", col="lightcoral")
boxplot(mtcars$qsec, main="Boxplot for 1/4 Mile Time", ylab="1/4 Mile Time", col="lightyellow")

# Identify outliers for each variable using the boxplot.stats() function
mpg_outliers <- boxplot.stats(mtcars$mpg)$out
hp_outliers <- boxplot.stats(mtcars$hp)$out
wt_outliers <- boxplot.stats(mtcars$wt)$out
qsec_outliers <- boxplot.stats(mtcars$qsec)$out

# Print the outliers
cat("Outliers for MPG:", mpg_outliers, "\n")
cat("Outliers for Horsepower:", hp_outliers, "\n")
cat("Outliers for Weight:", wt_outliers, "\n")
cat("Outliers for 1/4 Mile Time:", qsec_outliers, "\n")
```

Bonus 1: Visualizing Relationships Between Multiple Variables

```{r}
# Pair plot (scatterplot matrix) for mpg, hp, wt, qsec
pairs(mtcars[, c("mpg", "hp", "wt", "qsec")], main="Pair Plot of mpg, hp, wt, and qsec")
```

Bonus 2: Distribution Analysis

```{r}
# Density plots for mpg, hp, and wt
par(mfrow=c(1,3))  # Set up the plotting area to show 3 plots
plot(density(mtcars$mpg), main="Density Plot of MPG", xlab="Miles per Gallon")
plot(density(mtcars$hp), main="Density Plot of Horsepower", xlab="Horsepower")
plot(density(mtcars$wt), main="Density Plot of Weight", xlab="Weight")
```

Bonus 3: Identifying Multicollinearity

```{r}
# Correlation matrix
cor_matrix <- cor(mtcars[, sapply(mtcars, is.numeric)])
print(cor_matrix)
```

Bonus 4: Creating a Heatmap of Correlations

```{r}
library(ggplot2)
library(reshape2)
cor_matrix_melted <- melt(cor_matrix)
ggplot(cor_matrix_melted, aes(Var1, Var2, fill=value)) +
  geom_tile() + 
  scale_fill_gradient2(midpoint=0, low="blue", high="red", mid="white") + 
  theme_minimal() + 
  ggtitle("Heatmap of Correlation Matrix")
```

Bonus 5: Identifying Trends by Cylinder Group

```{r}
# Boxplot to compare mpg by cyl
boxplot(mpg ~ cyl, data=mtcars, main="MPG by Number of Cylinders", xlab="Number of Cylinders", ylab="Miles per Gallon", col="lightgreen")
```
